### 2020-08-26 paper/coloringBergeGraphs.tex
Lovasz Theta number

\subsection{Ellipsoid method}
\label{sec:ellipsoidMethod}

\subsubsection{Background}

\begin{wrapfigure}{r}{0.35\textwidth}
  \label{fig:c5}
	\input{tikzpictures/c5.tex}
	\caption{Example noisy channel}%
	% \vspace{-0.5cm}
	% \end{floatingfigure}
\end{wrapfigure}

Let us first discuss the background in which perfect graphs first arose. Suppose we have a noisy communication channel in which certain signal values can be confused with others. For instance, suppose our channel has five discrete signal values, represented as 0, 1, 2, 3, 4. However, each value of $i$ when sent across the channel can be confused with value $(i \pm 1)$ mod $5$. This situation can be modeled by a graph $C_5$ (\cref{fig:c5}) with in which vertices correspond to signal values and two vertices are connected iff values they represent can be confused.

We are interested in transission without possibility of confusion. In other words, we are interested in the Shannon capacity of the channel modeled by a graph \cite{Lovasz1979}. For this example it is possible for two values to be transmited without ambiguity e.g. values 1 and 4, which allows us to communicate $2^n$ bits of information in $n$ steps. But we could do better, for example we could communicate five two-step codewords e.g. "00", "12", "24", "43", "31". Each pair of these codewords includes at least one position where its values differ by two or more modulo 5, which allows the recipent to distinguish them without confusion.  This allows us to send $5^{n / 2}$ bits of information in $n$ steps, or effectively send $(5^{n/2})^{1/n} = \sqrt{5}$ bits per step.

Such methods provide a lower bound on the Shannon capacity of a graph

\subsubsection{\Lovasz number}
\TODO{background, definition}

@@

\subsubsection{Computing \boldmath$\vartheta$}
\TODO{TODO}

@@

\subsubsection{Stable set intersecting all maximum cardinality cliques}
Next, let us show how to find a stable set intersecting all the maximum cardinality cliques of $G$.

\begin{alg}
  \label{alg:ssIntersectingCliques}
  Input: A perfect graph $G = (V, E)$.

>>>>

\subsection{Information theory background}
\label{sec:ellipsoidMethod}

\subsubsection{Shannon Capacity of a graph}

The polynomial technique of coloring perfect graphs known so far arose in the field of semidefinite programming. Semidefinite programs are linear programs over the cones of semi-definite matrice. The connection of coloring graphs and the cones of semi-definite matrices might be surprising, so let us take a brief digression into the field of information theory, where we will see the connection more clearly. Also, this is the background which motivated Berge to introduce perfect graphs \cite{Chudnovsky2003}.

\begin{wrapfigure}{r}{0.35\textwidth}
  \label{fig:c5}
	\input{tikzpictures/c5.tex}
	\caption{Example noisy channel}%
	% \vspace{-0.5cm}
\end{wrapfigure}

Suppose we have a noisy communication channel in which certain signal values can be confused with others. For instance, suppose our channel has five discrete signal values, represented as 0, 1, 2, 3, 4. However, each value of $i$ when sent across the channel can be confused with value $(i \pm 1)$ mod $5$. This situation can be modeled by a graph $C_5$ (\cref{fig:c5}) with in which vertices correspond to signal values and two vertices are connected iff values they represent can be confused.

We are interested in transmission without possibility of confusion. For this example it is possible for two values to be transmited without ambiguity e.g. values 1 and 4, which allows us to send $2^n$ non-confoundable messages in $n$ steps. But we could do better, for example we could communicate five two-step codewords e.g. "00", "12", "24", "43", "31". Each pair of these codewords includes at least one position where its values differ by two or more modulo 5, which allows the recipent to distinguish them without confusion.  This allows us to send $5^{n / 2}$ non-confoundable messages in $n$ steps.

Let us be more precise. Given a graph $G$ modeling a communication channel and a number $k \geq 1$ we say that two messages $v_1v_2\ldots v_k$, $w_1w_2\ldots w_k \in G(V)^k$ of lenght $k$ are non-confoundable iff there is $1 \leq i \leq k$ such that $v_i$, $w_i$ are non-confoundable. We are interested in the maximum rate at which we can reliably transmit information (the \emph{Shannon capacity} of the channel defined by $G$) For $k = 1$, maximum number of messages we can send without confusion in a single step is equal to $\alpha(G)$.

To describe longer messages we indroduce \emph{Strong Product} $G \cdot H$ of two graphs $G = (V, E)$, $H = (W, F)$ as the graph with $V(G \cdot H) = V \times W$, with $(i, u)(j, v) \in E(G \cdot H)$ iff $ij \in E$ and $uv \in F$, or $ij \in E$ and $u = v$, or $i = j$ and $uv \in F$. Given channel modeled by $G$ it is easy to see that the maximum number of distinguishable words of length 2 is equal to $\alpha(G \cdot G)$, and in general the number of distinguishable words of length $k$ is equal to $\alpha(G^k)$ -- which gives us $\sqrt[k]{\alpha(G^k)}$ as the number of distinguishable signals per single transmission. So, we can define the Shannon capacity of the channel defined by $G$ as $\Theta(G) = \sup\limits_k \sqrt[k]{\alpha(G^k)}$.

Unfortunately, it is not known whether $\Theta(G)$ can be computed for all graphs (polynomial or not). If we could calculate $\alpha(G^k)$ for a first few values of $k$ (we will show how to do it in \cref{alg:maxStableSet}) we could have a lower bound on $\Theta(G)$. Let us now turn into search for some usable upper bound.

\subsubsection{\Lovasz number}

The following introduction of what has become known as \emph{\Lovasz number} is based on excellent lecture notes by \Lovasz \cite{Lovasz95}. \todo{rewrite this so its not blatantly copied.}

For a channel defined by graph $C_5$ using five messages of length 2 gives us a lower bound on $\Theta(G)$ equal $\sqrt{5}$ (as does calculating $\alpha(C_5^2)$).

Consider an "umbrella" in $\mathbb{R}^3$ with the unit vector $e_1 = (1, 0, 0)$ as its "handle" and 5 "ribs" of unit length \todo{picture}. Open it up to the point where non-consecutive ribs are orthogonal, that is form an angle of 90°. This way we get a representation of $C_5$ by 5 unit vectors $u_1, \ldots u_5$ so that each $u_i$ forms the same angle with $e_1$ and any two non-adjacent nodes are represented with orthogonal vectors. We can calculate $e_1^\intercal u_i = 5 ^ {-1/4}$.

It turns out, that we can obtain a similar representation of the nodes of $C_5^k$ by unit vectors $v_i \in \mathbb{R}^3k$, so that any two non-adjacent nodes are labeled with orthogonal vectors (this representation is sometimes called the \emph{orthogonal representation} \cite{Lovsz1989Orthogonal}). Moreover, we still get $e_1^\intercal v_i = 5^{-k/4}$ for every $i \in V(C_5^k)$ (the proof is quite technical and we omit it here).

If $S$ is any stable set in $C_5^k$, then $\{v_i, i \in S\}$ is a set of mutually orthogonal unit vectors so we get 
$$\sum\limits_{i\in S}(e_1^\intercal v_i)^2 \leq |e_1|^2 = 1$$ 
(if $v_i$ formed a basis then this inequality would be an equality).

On the other hand each term on the left hand side is $5^{-1/4}$, so the left hand side is equal to $|S|5^{-k/2}$, and so $|S| \leq 5^{k/2}$. Thus we get $\alpha(C_5^k) \leq 5 ^{k/2}$ and $\Theta(C_5) = \sqrt{5}$.

This method extends to any graph $G$ in place of $C_5$. All we have to do is find a orthogonal representation that will give us the best bound. So, we can define the \emph{\Lovasz number} of a graph $G$ as \todo{this equation does not really follow from the thought process above, it is a slightly different definition}: 
$$\vartheta(G) = \min\limits_{c,U} \max\limits_{i\in V} \frac{1}{(c^\intercal u_i)^2},$$

where $c$ is a unit vector in $\mathbb{R}^{|V(G)|}$ and $U$ is a orthogonal representation of $G$. But how can we construct an optimum (or even good) orthogonal representation? It turns out that it can be computed in polynomial time using semidefinite optimization.

\subsection{Computing \boldmath$\vartheta$}

\TODO{TODO}

\subsection{Coloring perfect graph using ellipsoid method}

@@

\TODO{TODO}

@@

\subsubsection{Stable set intersecting all maximum cardinality cliques}
Next, let us show how to find a stable set intersecting all the maximum cardinality cliques of $G$. 
\begin{alg}
  \label{alg:ssIntersectingCliques}
  Input: A perfect graph $G = (V, E)$.

### 2020-08-26 paper/definitions.tex
Amenable holes wip
Coloring
Lovasz Theta number

@@

An edge $\{x, y\}$ is said to \emph{join}, or be between vertices $x$ and $y$ and is denoted by $xy$. Thus $xy$ and $yx$ mean the same edge (all our graphs are \emph{unordered}). If $xy \in E(G)$ then $x$ and $y$ are adjacent, connected or neighboring. If $xy \notin E(G)$ then $xy$ is a \emph{nonedge} and $x$ and $y$ are \emph{anticonnected}.

@@

>>>>

\TODO{Run proofreader on all text (temporarily disabled because it slowed down IDE)}

@@

An edge $\{x, y\}$ is said to \emph{join}, or be between vertices $x$ and $y$ and is denoted by $xy$. Thus $xy$ and $yx$ mean the same edge (all our graphs are \emph{unordered}). If $xy \in E(G)$ then $x$ and $y$ are adjacent, connected or neighboring. By $N(x)$ we will denote the \emph{neighborhood} of $x$, that is all vertices $y$ such that $xy$ is an edge. If $xy \notin E(G)$ then $xy$ is a \emph{nonedge} and $x$ and $y$ are \emph{anticonnected}.

@@

\begin{defn}[stability number]
  A \emph{stability number} of a graph $G$, denoted as $\alpha(G)$, is a cardinality of its largest induced stable set.
\end{defn}
